---
title: "Twitter"
author: "Noah Love"
date: "2/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(readr)
df <- read_csv("A:\\Documents\\Semester 6\\data-mining-at-columbia\\My Class Notes\\small_non_retweets_dc_inaug_steal.csv")
View(df)


```


### Demo of different resampling methods

There are 3 types of resampling "things" that go on. 1) upsampling, just repeat positive cases that are rare. 2) Bootstrap, create copies by sampling with replacement. 3) Cross validation where are are separating part of the set. 

A news article pointed out that many of the plans for the seige of the capital occured on Twitter. So lets look at the word steal, often meaning "stop the steal"

```{r}
df$steal_flag <- as.numeric(df$steal_flag == "True")
#head(df[df$steal_flag > 0, "text"], 4)

mean(df$steal_flag)
pos_ind <- which(df$steal_flag > 0)
N <- length(pos_ind)*10
sample_ind <- sample(pos_ind, N, replace = TRUE)

#df is NOT overwritten!
new_df <- rbind(df[sample_ind, ],df)
ols <- lm(steal_flag ~ trump + capitol.police, data=new_df)
```

-Up-sampling is a hack to make the objective function give more weight to the positive yet rare cases. 


### Bootstrapping

```{r}
B <- 100

coeffs <- matrix(NA, nrow=B, ncol=3)
for(i in seq_len(B)){
  boot_ind <- sample(nrow(df), nrow(df), replace=TRUE)
  boot_df <- df[boot_ind, ]
  boot_ols <- lm(steal_flag ~ trump + capitol.police, data=boot_df)
  coeffs[i, ] <- boot_ols$coefficients
}

```


```{r}
k <- 5
# Make sure each split, has roughly the same number of records!
test_inds <- rep(1:k, nrow(df) / k)
# Make sure things are random, no bias!
test_inds <- sample(test_inds)

i <- 1
metrics <- c()
for (i in seq_len(k)) {
  is_test <- test_inds == i
  is_train <- !is_test
  ols <- lm(steal_flag ~ trump + capitol.police,
            data = df[is_train,])
  test_preds <- predict(ols, newdata = df[is_test, ])
  metrics[i] <- sum(abs(df$steal_flag[is_test] - test_preds))
}



```

