---
title: "Mapping practice"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tigris)
library(dplyr)
library(leaflet)
library(sp)
library(ggmap)
library(maptools)
library(broom)
library(httr)
library(rgdal)
```

```{r}
register_google(key = "AIzaSyAEkvFUTrN-n96QqRvpHxmtNyucQveP8oU")
```

 
```{r}
nyc_map <- get_map(location = c(lon = -74.00, lat = 40.71), maptype = "terrain", zoom = 11)
```


```{r}
ggmap(nyc_map)
```


Or make interactive maps using Leaflet for R in Rstudio.




```{r}
leaflet() %>%
  addTiles() %>%
  setView(-74.00, 40.71, zoom = 12)
```

Different Style in Leaflet

```{r}
leaflet() %>%
  addTiles() %>%
  setView(-74.00, 40.71, zoom = 12) %>%
  addProviderTiles("CartoDB.Positron")
```

Tigris can be used to grab census shapefiles at various levels.

```{r}
lookup_code("New York", "New York")
```
Then we download the data and take a look at it. The summary below shows the data associated with each polygon (e.g., the GEOID and NAME). We can access data columns via nyc_tracts@data.

```{r}
nyc_tracts <- tracts(state = '36', county = c('061','047','081','005','085'))

summary(nyc_tracts)
```

```{r}
plot(nyc_tracts)
```


Neighborhoods
Census tracts are great, but they might be too detailed for many purposes, where neighborhoods might be a more familiar unit of analysis. The PUMA Census neighborhoods for New York City lump together a few distinct areas (e.g., Chelsea, Flatiron, and the West Village), but Pediacities has more fine-grained neighborhood shapefiles available as a GeoJSON file.

```{r}
r <- GET('http://data.beta.nyc//dataset/0ff93d2d-90ba-457c-9f7e-39e47bf2ac5f/resource/35dd04fb-81b3-479b-a074-a27a37888ce7/download/d085e2f8d0b54d4590b1e7d1f35594c1pediacitiesnycneighborhoods.geojson')
nyc_neighborhoods <- readOGR(content(r,'text'), 'OGRGeoJSON', verbose = F)
```


```{r}
summary(nyc_neighborhoods)
```
ggplot
ggplot can plot shapefiles but requires them to be converted from a spatial polygon data frame format to plain old data frames first. broom handles this with its tidy command. (fortify from the sp package also works.)

```{r}
nyc_neighborhoods_df <- tidy(nyc_neighborhoods)
```


```{r}
ggplot() + 
  geom_polygon(data=nyc_neighborhoods_df, aes(x=long, y=lat, group=group))
```

Overlay the two!


```{r}
ggmap(nyc_map) + 
  geom_polygon(data=nyc_neighborhoods_df, aes(x=long, y=lat, group=group), color="blue", fill=NA)
```

Lets make it interactive!
```{r}
leaflet(nyc_neighborhoods) %>%
  addTiles() %>% 
  addPolygons(popup = ~neighborhood) %>%
  addProviderTiles("CartoDB.Positron")
```

